---
title: "Option Pricing"
author: "Sanjiv Das"
date: "7/28/2018"
output: html_document
---

## Black-Merton-Scholes (1973)

The famous Black-Scholes (1973) and Merton (1973) option pricing model is used to price European call and put options. The equation for the call price is as follows:

$$
C = S e^{-qT} N(d_1) - K e^{-rT} N(d_2)
$$
where 

$$
d_1 = \frac{\ln(S/K)+(r-q+\frac{1}{2}\sigma^2)T}{\sigma \sqrt{T}}; \quad d_2 = d_1 - \sigma \sqrt{T}
$$

The corresponding equation for the put price is:

$$
P = K e^{-rT} N(-d_2) - S e^{-qT} N(-d_1)
$$

The code for the model is here:

```{r}
#Black-Merton-Scholes 1973
#sig: volatility
#S: stock price
#K: strike price
#T: maturity
#r: risk free rate
#q: dividend rate
#cp = 1 for calls and -1 for puts
#optprice: observed option price
bms73 = function(sig,S,K,T,r,q,cp=1) {
    d1 = (log(S/K)+(r-q+0.5*sig^2)*T)/(sig*sqrt(T))
    d2 = d1 - sig*sqrt(T)
    if (cp==1) {
        optval = S*exp(-q*T)*pnorm(d1)-K*exp(-r*T)*pnorm(d2)
        }
    else {
        optval = -S*exp(-q*T)*pnorm(-d1)+K*exp(-r*T)*pnorm(-d2)		
    }
}
```

We run example prices using some sample parameters. 

```{r}
S=50; K=51; T=1; sig=0.15; rf=0.02; q=0.01
cat("Call price: "); print(bms73(sig,S,K,T,rf,q))
cat("Put price: "); print(bms73(sig,S,K,T,rf,q,cp=-1))
```

## Pricing by Simulation

We use a parallelized approach:

```{r}
m = 50000  #No of simulation paths
n = 50     #No of time periods
h = T/n    #Time interval
stk = matrix(0,m,n+1)  #Grid for stock paths
e = matrix(rnorm(m*n),m,n)
stk[,1] = S
for (t in 2:(n+1)) {
  stk[,t] = stk[,t-1]*exp((rf-q-0.5*sig^2)*h + sig*sqrt(h)*e[,t-1]) #vectorization on columns (time)
}
c = mean(pmax(0,stk[,n+1]-K))*exp(-rf*T)
p = mean(pmax(0,K-stk[,n+1]))*exp(-rf*T)
print(c(c,p))
```

## Antithetic Variance Reduction

By choosing opposing complementary random variables, we can reduce the variance of the estimates. 

```{r}
m = 50000  #No of simulation paths
n = 50     #No of time periods
h = T/n    #Time interval
stk = matrix(0,m,n+1)  #Grid for stock paths

#Antithetic
e1 = matrix(rnorm(m/2*n),m/2,n)
e2 = -e1
e = rbind(e1,e2)

stk[,1] = S
for (t in 2:(n+1)) {
  stk[,t] = stk[,t-1]*exp((rf-q-0.5*sig^2)*h + sig*sqrt(h)*e[,t-1])
}
c = mean(pmax(0,stk[,n+1]-K))*exp(-rf*T)
p = mean(pmax(0,K-stk[,n+1]))*exp(-rf*T)
print(c(c,p))
```

## Longstaff-Schwartz Model

```{r}
LSM = function(S0,K,sig,r,h,n,m) {
  stock = stk  # Initialize stock grid using previous grid
  option = matrix(0,m,n+1)  # Initialize option grid
  option[,n+1] = pmax(0, K - stock[,n+1]) # option at maturity
  # Backward recursion
  for (i in seq(n, 2, -1)) { 
    index = which(K - stock[, i] > 0)  #rows early exercise possible
    W = exp(-r * h) * option[, i + 1] # continuation value 
    Y = W[index] # Vector of continuation values
    S = stock[index, i] # stock price vector for regression 
    # matrix for 3rd order polynomial regression 
    S = matrix(c(S, S^2, S^3), length(S), 3); 
    beta = lm(Y ~ S); 
    # continuation value step 
    earlyIndex = matrix(0, m, 1)
    earlyIndex[index] = 1
    S = stock[, i]
    S = matrix(c(S, S^2, S^3), length(S), 3)
    predictedY = S %*% beta$coefficients[2:4] + beta$coefficients[1]
    contVal = predictedY * earlyIndex
    earlyIndex2 = matrix(0, m, 1)
    index = which(K - stock[, i] > contVal)
    earlyIndex2[index] = 1
    earlyIndex = earlyIndex * earlyIndex2
    option[, i] = earlyIndex * (K - stock[, i]) + (1 - earlyIndex) * W
  } 
  # Complete option value at initial date
  option[, 1] = option[, 2] * exp(-r * h); 
  result = max(mean(option[, 1]), K - S0); 
}

AmPut = LSM(S,K,sig,rf,h,n,m)
print(AmPut)
```

